\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-02-15 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
making corrections

\medskip
Out of this mixed and complex various geological formations, multiphase flows through porous media. play the most important role in understanding the afformentioned operations. These often involves the displacement of a non-wetting invading fluid from a porous medium by a wetting fluid. Modeling of multiphase flow, on the other hand is still an enormous technical and computational challenge. In all branches of fluid mechanics, there are the instabilities, also know as flow instabilities. The first distinction should be made between microscopic and macroscopic instabilities. The term microscopic instabilities is used for the phenomena occurring locally; for example, the Helmholtz and Taylor instabilities or bubble collapse. On the other hand the macroscopic instabilities involve, e.g.the entire two-phase flow system. In this paper, the main focus is kept on macroscopic phenomena. Part of flow instabilities is the viscous flow instabilities also know as \textit{viscous fingering}. In large scale field operations it is the condition whereby the interface of two fluids, such as oil and water, bypasses sections of reservoir as it moves along, creating an uneven, or fingered, profile. Fingering is a relatively common condition in reservoirs with water-injection wells. The result of fingering is an inefficient sweeping action that can bypass significant volumes of recoverable oil and, in severe cases, an early breakthrough of water into adjacent production wellbores. It can occur both within a single phase or when multiple phases flow through porous media.

@@

\medskip
The objective of this study is to investigate and accurately describe the behaviour of fingering in multiphase ($2$-phase) flows using a control volume finite element method (CVFEM). Considering the limitations of finite differences formulations the CVFE the model formulation will use a conservative computational multi-fluid porous media flow model able to exploit mesh adaptivity methods on a fully-unstructured tetrahedral grid. Also two key characteristics that describe the model above are: (a) the hybrid element pair $P_{n}DGP_{n+1}$ which provides discontinuous and piecewise representation for velocity and pressure fields at FE space and (b) the overlapping control volume finite element method (CVFEM)formulation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing equations and numerical scheme}\label{equations_scheme}      
Multiphase flow in porous media is described by mass conversation (or transport) equation and Darcy velovities for each of the phase $\alpha$ ($\alpha$ = g, o, w, for gas, oil, and water phases).
Fingering behaviour is expected to be different from single-phase (piston shape front) due to the effect of different peremeabilities. This will change the mobility contrast between two phases for a given \textit{viscosity} ratio. Because of the  different saturation levels among different layers, eventually this will enhance the flow instabilities. In this paper, the test-cases are also based on the discontinuous Galerkin (DG) method to simulate viscous fingering. Roughly speaking, DG-method combines advantages of classical finite volume and finite element methods.

>>>>

\medskip
Out of this mixed and complex various geological formations, multiphase flows through porous media. play the most important role in understanding the afformentioned operations. These often involves the displacement of a non-wetting invading fluid from a porous medium by a wetting fluid. Modeling of multiphase flow, on the other hand is still an enormous technical and computational challenge. In all branches of fluid mechanics, there are the instabilities, also know as flow instabilities. The first distinction should be made between microscopic and macroscopic instabilities. The term microscopic instabilities is used for the phenomena occurring locally; for example, the Helmholtz and Taylor instabilities or bubble collapse. On the other hand the macroscopic instabilities involve, e.g.the entire two-phase flow system. In this paper, the main focus is kept on macroscopic phenomena. Part of flow instabilities is the viscous flow instabilities, also know as \textit{viscous fingering}. In large scale field operations it is the condition whereby the interface of two fluids, such as oil and water, bypasses sections of reservoir as it moves along, creating an uneven, or fingered, profile. Fingering is a relatively common condition in reservoirs with water-injection wells. The result of fingering is an inefficient sweeping action that can bypass significant volumes of recoverable oil and, in severe cases, an early breakthrough of water into adjacent production wellbores. It can occur both within a single phase or when multiple phases flow through porous media. This instability is essentially governed by the mobility-ratio, MR, which is defined as the ratio of the mobility of the displacing fluid over that of the displaced fluid. In a water-oil system the MR is given by eq.\ref{eq:MR}    However, other important factors also influence the severity of viscous fingering: heterogeneous permeability fields, gravity, anisotropic dispersion, velocity dependence of dispersion, nonmonotonic viscosity profiles, etc. High mobility-ratio displacement processes present very  complex  fingering  patterns.

\begin{equation}
 MR  = \frac{k_{rw} \; \mu_{w}}{k_{ro} \; \mu_{o}} 
\label{eq:MR}
\end{equation}

@@

\medskip
The objective of this study is to investigate and accurately describe the behaviour of fingering in multiphase ($2$-phase) flows using a control volume finite element method (CVFEM). Considering the limitations of finite differences formulations the CVFE the model formulation will use a conservative computational multi-fluid porous media flow model able to exploit mesh adaptivity methods on a fully-unstructured tetrahedral grid. Also two key characteristics that describe the research concept above are: (a) the hybrid element pair $P_{n}DGP_{n+1}$ which provides discontinuous and piecewise representation for velocity and pressure fields at FE space and (b) the overlapping control volume finite element method (CVFEM)formulation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing equations and numerical scheme}\label{equations_scheme}      
Multiphase flow in porous media is described by mass conversation and Darcy equations for all the phase $\alpha$ ($\alpha$ = g, o, w, for gas, oil, and water phases).
Fingering behaviour is expected to be different from single-phase (piston shape front) due to the effect of different peremeabilities. This will change the mobility contrast between two phases for a given \textit{viscosity} ratio. Because of the  different saturation levels among different layers, eventually this will enhance the flow instabilities. In this paper, the test-cases are also based on the discontinuous Galerkin (DG) method to simulate viscous fingering. Roughly speaking, DG-method combines advantages of classical finite volume and finite element methods.
